---
title: "HW4"
author: "Burak Can Helvacı"
---
## Task : Multiple Instance Learning
Suggest two alternative bag-level representations for the given multiple instance learning problem. Based on the proposed bag-level representations for Musk1 dataset, evaluate at least two reasonable classifiers (of your choice). Specify the best set of parameter combination for the proposed representations and your classifier. Use the accuracy based on 10-fold cross-validation on the training data as your primary performance metric in your evaluations.

Alternative Representations:
*Euclidian Distance
*Principal Component Analysis

Alternative Classifiers
*Random Forest
*Gradient Boosting

```{r}
library(data.table)
library(caret)
library(ranger)
library(gbm)

musk <- read.csv("Musk1.csv", header = FALSE)
dim(musk)
head(musk)
```

As first representation I choose distance matrix of instances using euclidian distance to decide which instance to represent each bag.
The instance which is the most dissimilar to instances in other bags is chosen to represent the respective bag.
Dissimilarity is calculated as sum of distances to out of bag instances. 

```{r, error=TRUE}
musk2 <- musk
musk2 <- scale(musk2[,3:168])
musk2 <- as.matrix(dist(musk2))
musk2 <- cbind(musk$V1, musk$V2, 1:length(musk$V1),musk2)
musk2 <- as.data.frame(musk2)
setnames(musk2, c("V1","V2","V3"), c("BagClass","BagId","Instance_Id"))

lst <- c()
for(i in 1:length(musk2$BagId)){
  aa = musk2[musk2$BagId == i,]
  aa = aa[,-1*(aa$Instance_Id +3)]
  bb = rowSums(aa[,4:ncol(aa)])
  cc = which.max(bb)
  lst <- c(lst,aa$Instance_Id[cc])
}

musk3 <- musk[lst,]
musk3 <- musk3[,-2]
setnames(musk3, c("V1"), c("class"))
```

As second representation I used PCA. For every class, PCA is carried out. In total there are 92 separate PCA is conducted.
First component (eigenvector) of the PCA is used to represent the class.

```{r}
lst2 = list()
musk4 <- musk
musk4 <- setnames(musk, c("V1","V2"), c("BagClass","BagId")) 

for (i in 1:length(unique(musk4[["BagId"]]))){
  res = prcomp(musk4[musk4$BagId==i, 3:ncol(musk4)])
  lst2[[i]] = t(res$rotation[,1])
}

musk4 <- do.call(rbind.data.frame, lst2)
musk4$class <- c(rep(1,47),rep(0,45))
```

As first classifier. I will use random forest. 
Random forest with first representation
```{r}
n_folds <- 10
fitControl <- trainControl(method = "cv", number = n_folds,
                           classProbs=TRUE, summaryFunction=twoClassSummary)

rf_grid <- expand.grid(mtry = seq(10, 30, 2),
                       splitrule = c("gini", "extratrees"),
                       min.node.size = c(5))

musk3$class <- ifelse(musk3$class==1, "Yes", "No")
setnames(musk3, "class", "y")  

set.seed(1)                      
rf_fit1 <- caret::train(y ~ ., data = musk3,
                method = "ranger", metric="ROC",
                trControl = fitControl, 
                tuneGrid = rf_grid)

rf_fit1$bestTune
plot(rf_fit1)
rf_fit1$results[20,]
preds <- predict(rf_fit1)
cf <- table(preds,musk3$y)
accuracy_rf = sum(diag(cf)/sum(cf))
cf
accuracy_rf
```

Random forest with second representation

```{r}
musk4$class <- ifelse(musk4$class==1, "Yes", "No")
setnames(musk4, "class", "y") 

set.seed(1)                      
rf_fit2 <- caret::train(y ~ ., data = musk4,
                        method = "ranger", metric="ROC",
                        trControl = fitControl, 
                        tuneGrid = rf_grid)
rf_fit2$bestTune
rf_fit2$results[12,]
plot(rf_fit2)
preds <- predict(rf_fit2)
cf <- table(preds,musk4$y)
accuracy_rf = sum(diag(cf)/sum(cf))
print(cf)
print(accuracy_rf)
```

As last classifier. I will use gradient boosting.
Gradient boosting with first representation

```{r}
gbmGrid <- expand.grid(interaction.depth = c(3, 4, 6, 8), 
                       n.trees = c(1:5)*100, 
                       shrinkage = seq(0.02, 0.1, 0.02),
                       n.minobsinnode = 5)
set.seed(1)                        
gbm_fit1 <- train(y ~ ., data = musk3, 
                 method = "gbm", 
                 trControl = fitControl, metric='ROC',
                 tuneGrid = gbmGrid,
                 verbose=F)

gbm_fit1$bestTune
gbm_fit1$results[90,]
plot(gbm_fit1)
preds <- predict(gbm_fit1)
cf <- table(preds,musk3$y)
accuracy_rf = sum(diag(cf)/sum(cf))
print(cf)
print(accuracy_rf)
```

Gradient boosting with second representation

```{r}
set.seed(1)                        
gbm_fit2 <- train(y ~ ., data = musk4, 
                  method = "gbm", 
                  trControl = fitControl, metric='ROC',
                  tuneGrid = gbmGrid,
                  verbose=F)

gbm_fit2$bestTune
gbm_fit2$results[69,]
plot(gbm_fit2)
preds <- predict(gbm_fit2)
cf <- table(preds,musk4$y)
accuracy_rf = sum(diag(cf)/sum(cf))
print(cf)
print(accuracy_rf)
```

## Conclusion
Random forest gives better results than gradient boosting.
Euclidian distance representation gives better results than principal component representation.


